High-resolution ALMA observations of V4046\,Sgr: a circumbinary disc with a thin ring

### 2020-10-13 Draft_V4046Sgr_mnras.tex
with SC edits

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[R. Martinez Brunner et al.]{
R. Martinez-Brunner,$^{1}$\thanks{E-mail: rmartinezbrunner@gmail.com}
A. N. Other,$^{2}$
Third Author$^{2,3}$
and Fourth Author$^{3}$
\\
% List of institutions
$^{1}$Chile\\
$^{2}$Department, Institution, Street Address, City Postal Code, Country\\
$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country
}

@@

% Abstract of the paper
\begin{abstract}
    V4046 Sgr is a nearby pre-main sequence spectroscopic binary system that hosts a gas-rich disc known for its wide gaps and dusty ring as seen by the Atacama Large Millimeter/submillimetre Array (ALMA). We present new high resolution ($\sim$20\, mas) ALMA observations of the 1.3\,mm radio continuum. These observations, combined with SPHERE-IRDIS polarized scattered light images, allow us to characterize the disc structure. We reproduce the available observations, including the spectral energy distribution, in terms of a physical structure using a parametric radiative transfer (RT) model. We report a two-ringed structure for the 1.3 mm dust emission with a narrow, previously unseen, less than 4.17$\pm$0.94\,au-wide inner ring centred at 13.46$\pm$0.43\,au. The ring is surrounded by a $\sim$9\,au-wide gap, flanked by a very sharp outer ring starting at 24\,au. We deproject the images in polar coordinates to better constrain the disc radial structure. The radial profile of the outer ring exhibits an intriguing discontinuity at $\sim$35\,au. The model predicts a three-ringed structure for the small-dust population with an inner ring at $\sim$6\,au hiding under the coronagraph in the SPHERE observations. The shadow of the secondary star reported previously from scattered-light data is also reproduced by our RT model. The polarized light image reveals strong forward-scattering at 1.65\,$\micron$, which we reproduced with dust grains of no less than 0.4\,$\micron$, hinting at a deficit of very-small-grains. Using the parametric model scale-height as an estimated height, we found that the thin inner ring is stable. Deeper ALMA observations may confirm the lopsided 6\,au ring.
\end{abstract}

@@

Recent observations of circumstellar discs have revolutionised the way we think about protoplanetary discs, but the focus of resolved imaging with the Atacama Large Millimeter/submillimetre Array (ALMA) or with VLT/SPHERE has been mainly been towards the brighter sources. ***EXPAND, I would give a general description on substructures and cite ``e.g. Andrews (2020) for a recent review''***

V4046 Sagittarii (Sgr) is is a double-lined spectroscopic binary of K-type stars (K5 and K7) with very similar masses of 0.90$\pm$0.05\,M$_{\sun}$ and 0.85$\pm$0.04\,M$_{\sun}$ \citep{Rosenfeld_2012} on a close ($a \approx 0.041$\,au), circular ($e\leq0.001$) orbit, with an orbital period of 2.42 days \citep{refId0}. The V4046\,Sgr system, located at 72.41$\pm$0.34\,pc \citep{Gaia}, is a member of the $\beta$ Pictoris moving group \citep{Zuckerman_2004}, with an estimated age of 23$\pm$3\,Myr \citep{Mamajek_2014}. The V4046\,Sgr binary system hosts a massive ($\sim$0.1\,M$_{\sun}$) circumbinary disc extending to $\sim$300\,au \citep{Rosenfeld_2013, Rodriguez_2010}. The disc has a vast variety of molecules with a detailed sub-arcsecond ALMA molecular line imaging study by \citet{Kastner_2018}.

The Disks ARound T Tauri Stars (DARTTS) program was first presented in \citet{Avenhaus_2018} with the aim to image eight stars with the Spectro-Polarimeter High-contrast Exoplanet REsearch (SPHERE). The sample is not biased towards exceptionally bright and large disks, and consists of only solar-mass stars. A second part of the survey increased the number of sources by presenting 21 new images of circumstellar disks \citep{Garufi2020}. The observations revealed diversed structures and morphologies in the scattering surface of these disks. This letter is part of our new DARTTS survey with ALMA (DARTTS-A) which will present millimeter observations of nine protoplanetary disks previously imaged in polarized scattered light in DARTTS-S.

Here, we present new 1.3\,mm continuum ALMA observation of V4046\,Sgr. The structure of this letter is as follows. The new 1.3\,mm continuum observation is described alongside with a SPHERE-IRDIS \textit{H}-band image in Section \ref{sec:Observations}. The structure of the parametric model used to simulate the disc is presented in Section \ref{sec:model}. It is worth mentioning that previous models of V4046\,Sgr have been made \citep{Ru_z_Rodr_guez_2019, Rosenfeld_2013, 2019ApJ...882..160Q} but with the difference that our new reference observation of radio wavelengths has a sub-arcsecond resolution, allowing us to discover previously unseen underlying substructures of the large-dust grain disc. The results of the model and the discussion are commented Section \ref{sec:results} and the main conclusions are described in Section \ref{sec:Conclusions}.

@@

New ALMA observations of V4046\,Sgr were obtained in 2017 as part of the Cycle 5 program 2017.1.01167.S (PI: S. Perez). These were acquired in the context of a larger survey for 9 optically visible TTauri discs at $\sim$50\,mas resolution, the DARTTS-A program (Perez et al. {\em in prep}). The survey simultaneously mapped 1.3\,mm continuum and the J = 2$-$1 line of $^{12}$CO with the C43-8 array configuration in band 6 (211-275\,GHz). This work focuses exclusively on the data obtained in the 1.3\,mm continuum for our source, taken with a beam size of $0\farcs04 \,\times \, 0\farcs06$ (in natural weights).

In this investigation, we use the previous polarimetric image at 1.65\,$\micron$ part of the DARTTS-SPHERE I survey \citep{Avenhaus_2018}. The image was taken at the ESO Very Large Telescope with the SPHERE-IRDIS instrument in differential polarization imaging (DPI) mode, and a complete description of the data reduction was presented in \citet{Avenhaus_2018}.

For our analysis of the ALMA image we used an image reconstruction strategy, using the non-parametric image synthesis of the \textsc{uvmem} package \citep{2006ApJ...639..951C, 2018A&C....22...16C} it is possible to super-resolve the clean beam, obtaining an effective angular resolution $\sim$2-3 times higher than in the clean image with natural weights. The model obtains this higher resolution through fitting the data using Chi-squared minimisation. Here we adopted a pure $\chi^2$ model image. The resulting \textsc{uvmem} image shown in Fig. \ref{fig:two} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them. The wide and brighter outer ring has its peak intensity at $\sim$30\,au, has a surface brightness break at $\sim$35\,au, and reaches out to $\sim$60\,au. The thin and weak inner ring had never seen before, however \citet{Ru_z_Rodr_guez_2019} did anticipate its existence.

\begin{figure*}
  \includegraphics[width=\textwidth]{hot_two_E.pdf}
  \caption{Comparison of observations and simulated images at 1.65\,$\micron$ and 1.3 mm continuum of the circumbinary disc orbiting V4046\,Sgr. From top to bottom: DPI image and ALMA Band 6 observations; simulated images of the parametric model. Top left panel: SPHERE-IRDIS \textit{H}-band image with a yellow filled circle that illustrate the N\_ALC\_YJH\_S coronagraph (inner working angle $\sim0\farcs12$, or $\sim$8.6\,au at 72.4\,pc). Top right panel: 1.3\,mm continuum \textsc{uvmem} model image. The yellow ellipse shows the size of the natural-weighted beam: $ 0\farcs04 \, \times \, 0\farcs06$. The inset displays a enlarged view of the central emission, the green star marks the star positions. Bottom left panel: 1.65\,$\upmu m$ simulated image of the parametric model. Bottom right panel: 1.3\,mm simulated image of the parametric model. The yellow ellipse shows the size of the simulated beam: $ 0\farcs04 \, \times \, 0\farcs06$. For all the images in the figure the colour scale is linear.}
  \label{fig:two}
\end{figure*}

The inner ring is surprisingly narrow and seems to be off-centre relative to the GAIA stellar position (at the origin of coordinates in Fig. \ref{fig:two}). The ring width can be measured in polar coordinates following the determination of the inner ring centre and orientation. We minimised the dispersion of the disc radial profiles, from 6\,au to 19\,au (aiming for the inner ring) for a PA of 74.6\,$\degr$, an inclination of 33.9\,$\degr$ and a centre at $\Delta \alpha = 9\pm0.05$\,mas $\Delta \delta = 0.1\pm0.04$\,mas, relative to the stars.

Assuming a Gaussian dust inner ring, we obtained its mean width and location by finding for each azimuth the Gaussian intensity profile that best describes the ring. This gave us a FWHM of 4.17$\pm$0.94\,au and that it is located at 13.46$\pm$0.43\,au from the stars (See Fig. \ref{fig:polarring}). As the \textsc{uvmem} model image has an effective angular resolution of $\sim$1/3 that of the natural-weighted beam ($0\farcs04 \,\times \, 0\farcs06$), these measurements only gives an upper limit of the real width.

Doing one more time the minimisation of the dispersion of the disc radial profiles, but this time aiming for the outer ring from 20\,au to 70\,au, for a PA of 76.8\,$\degr$, with an inclination of 34.0\,$\degr$ and a centre at $\Delta \alpha = 19\pm0.04$\,mas $\Delta \delta = 7\pm0.03$\,mas relative to the stars, we can compare the measured location of the inner ring according with this parameters versus to the location obtained before. The deviation for each azimuth is shown in Fig. \ref{fig:polarring}, and is not trivial since it depend on a few parameters such as the different inclinations of the rings, the difference in position angle and the shift of the centres relative to the stars. But even though those parameters play an important role, a possible small difference in eccentricities between the inner and the outer ring could also help explain this deviation between the two measurements. ***

@@

The scattered-light image shown in Fig. \ref{fig:two} reveals the presence of a double ring structure in the micron-sized dust distribution. The observed morphology presents an inner cavity of $\sim$10\,au in radius and two rings located at 14.10$\pm$0.01 and 24.62$\pm$0.08\,au with a small gap between them at $\sim$20\,au. The observed second ring might correspond to the inner wall of the 1.3\,mm continuum emission outer ring \citep{Ru_z_Rodr_guez_2019}. Two other important features that are present in the image are: the brightness asymmetry and the shadows projected on the disc produced by the close binary system as they eclipse each other, described in detail by \citet{dOrazi}.

\citet{dOrazi} also measured the binary phase in the scattered-light observation around 106\,$\degr$, setting $\phi=$0 from the North and increasing clockwise. Using this measurement, the binary phase was calculated at the time of the ALMA observation at $\phi=$13\,$\degr$. There is no clear signal of any shadow similar to the ones present in the DPI image, this could be explained due to the slower cooling of larger grains (cite Casassus?***).

\begin{figure}
    \includegraphics[width=\columnwidth]{polar_ring_aprox_and_diff_inner.pdf}
    \caption{(a) Polar decomposition of the 1.3\,mm continuum image showing the centres (solid line) of Gaussian fits and the measured FWHM (blue region between the dotted lines) of the inner thin ring for each azimuthal angle. This was obtained centring the \textsc{uvmem} image optimising for the outer ring. (b) Shows the measured location of the inner ring, where the orange line represents the centres of the fitted Gaussian profile obtained in the first panel and blue line represents the same but obtained centring the \textsc{uvmem} image optimising for the inner ring.}
    \label{fig:polarring}
\end{figure}

@@

\section{Parametric radiative transfer model} \label{sec:model}

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[R. Martinez Brunner et al.]{
Rafael Martinez-Brunner,$^{1}$\thanks{E-mail: rmartinezbrunner@gmail.com}
Simon Casassus,$^{1}$
Sebasti\'an P\'erez,$^{2}$
et al. 
\\
% List of institutions
$^{1}$Departamento de Astronom\'{\i}a, Universidad de Chile, Casilla 36-D, Santiago, Chile\\
$^{2}$Universidad de Santiago de Chile, Av. Ecuador 3659, Santiago\\
}

@@

% Abstract of the paper
\begin{abstract}
  The nearby V4046\,Sgr spectroscopic binary  hosts a gas-rich disc known for its wide cavity and dusty ring. We present new high resolution ($\sim$20\, mas) ALMA observations of the 1.3\,mm  continuum. The comparison of these  observations, combined with SPHERE-IRDIS polarized  images and a well-sampled spectral energy distribution (SED), against radiative transfer (RT) predictions carried out with the RADMC3D package, allow us to propose a physical model for the source. The new ALMA data reveal a very fine ring at a radius of 13.46$\pm$0.43\,au (Ring13), with a marginally resolved radial width of  4.17$\pm$0.94\,au.  Ring13 is the brightest structure in scattered-light, where it is surrounded by a $\sim$9\,au-wide gap , and it is flanked by a very bright mm outer ring (Ring24) with a sharp inner edge at 24\,au. The steeply decreasing radial slope of Ring24 breaks at $\sim$35\,au into a shallow tail. The RT model requires an inner ring at $\sim$6\,au (Ring6) in small dust grains, hiding under the IRDIS coronagraph, and that surrounds an inner circumbinary disk. Faint mm-continuum coincident with Ring6 is picked up by ALMA, whose morphology  suggests that Ring6  is lopsided or shadowed by the secondary. The previously reported scattered-light shadow of the secondary star is  also reproduced by the RT model.  The surpising  thin Ring13 is nonetheless **10*** times wider than the model scale height, and could thus be long-lived.   The strong near-far disc asymmetry  at 1.65\,$\micron$ points at a very forward-scattering phase function, and requires  grain radii  of no less than  0.4\,$\micron$. 
\end{abstract}

@@

Recent observations of young circumstellar discs have transformed  current knowledge of planet formation, but the focus of resolved imaging with the Atacama Large Millimeter/submillimetre Array (ALMA) or with VLT/SPHERE has  mainly been towards the brighter sources \citep[see][for a review]{Andrews2020arXiv200105007A}. The Disks ARound T Tauri Stars (DARTTS) program was first presented in \citet{Avenhaus_2018} with the aim to image eight stars with the Spectro-Polarimeter High-contrast Exoplanet REsearch (SPHERE). The sample is not biased towards exceptionally bright and large disks, and consists of only solar-mass stars. A second part of the survey increased the number of sources by presenting 21 new images of circumstellar disks \citep{Garufi2020}. The observations revealed diversed structures and morphologies in the scattering surface of these disks. This letter on V4046 Sagittarii (Sgr) is part of our new DARTTS survey with ALMA (DARTTS-A) which will present millimeter observations of nine protoplanetary disks previously imaged in polarized scattered light in DARTTS-S.

V4046\,Sgr is a double-lined spectroscopic binary of K-type stars (K5 and K7) with very similar masses of 0.90$\pm$0.05\,M$_{\sun}$ and 0.85$\pm$0.04\,M$_{\sun}$ \citep{Rosenfeld_2012} on a close ($a \approx 0.041$\,au), circular ($e\leq0.001$) orbit, with an orbital period of 2.42 days \citep{refId0}. It is a member of the $\beta$ Pictoris moving group \citep{Zuckerman_2004}, with an estimated age of 23$\pm$3\,Myr \citep{Mamajek_2014}, and its distance is  72.41$\pm$0.34\,pc \citep{Gaia}. V4046\,Sgr hosts a massive ($\sim$0.1\,M$_{\sun}$) circumbinary disc extending to $\sim$300\,au \citep{Rosenfeld_2013, Rodriguez_2010}, with a rich observable chemical  diversity  \citep{Kastner_2018}.

The structure of this letter is as follows. The observations, including  new 1.3\,mm continuum data, are described in Section \ref{sec:Observations}. We interpret the data in terms of a parametric model,  presented in Section \ref{sec:model}. Previous models of V4046\,Sgr have been made \citep{Ru_z_Rodr_guez_2019, Rosenfeld_2013, 2019ApJ...882..160Q} but our new ALMA data bring additional information. Our results are discussed in  Section \ref{sec:results} and summarised in  Section \ref{sec:Conclusions}.

@@

New ALMA observations of V4046\,Sgr were obtained in 2017 as part of the Cycle 5 program 2017.1.01167.S (PI: S. Perez). These were acquired in the context of the DARTTS-A program (Perez et al. {\em in prep}), a larger survey for 9 optically visible TTauri discs at $\sim$50\,mas resolution. The survey simultaneously mapped the 1.3\,mm continuum and the J = 2$-$1 line of $^{12}$CO with the C43-8 array configuration in band 6 (211-275\,GHz). This work focuses exclusively on the data obtained in the 1.3\,mm continuum for our source, taken with a beam size of $0\farcs04 \,\times \, 0\farcs06$ (in natural weights). **Seba, sure that the survey only used C43-8?***

In this investigation, we use the  polarimetric image at 1.65\,$\micron$ previously published as part of the DARTTS-SPHERE I survey \citep{Avenhaus_2018}. The image was taken at the ESO Very Large Telescope with the SPHERE-IRDIS instrument in differential polarization imaging (DPI) mode, and a complete description of the data reduction was presented in \citet{Avenhaus_2018}.

For our analysis of the ALMA image we used an image reconstruction strategy, using the non-parametric image synthesis of the \textsc{uvmem} package \citep{2006ApJ...639..951C, 2018A&C....22...16C} it is possible to super-resolve the clean beam, obtaining an effective angular resolution $\sim$3 times finer than  natural weights. The model image is obtained by fitting the data using Chi-squared minimisation, with a measure of regularization when required. Here we adopted a pure $\chi^2$ model image ***sure about this? what is the name of the image you used?****. The resulting \textsc{uvmem} image shown in Fig. \ref{fig:two} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them, i.e. Ring13  at 13\,au and Ring24 starting at 24\,au. The wide and bright  Ring24  reaches its peak intensity at $\sim$30\,au, beyond which it drops steeply before breaking at  $\sim$35\,au into a  shallow tail. While this is the first observation of Ring13,  \citet{Ru_z_Rodr_guez_2019} did anticipate its existence ****HOW? EXPAND ON THIS***

\begin{figure*}
  \includegraphics[width=\textwidth]{hot_two_E.pdf}
  \caption{Comparison of observations and simulated images at 1.65\,$\micron$ and 1.3 mm continuum of the circumbinary disc orbiting V4046\,Sgr. From top to bottom: DPI image and ALMA Band 6 observations; simulated images of the parametric model. Top left panel: SPHERE-IRDIS \textit{H}-band image with a yellow filled circle that illustrate the N\_ALC\_YJH\_S coronagraph (inner working angle $\sim0\farcs12$, or $\sim$8.6\,au at 72.4\,pc). Top right panel: 1.3\,mm continuum \textsc{uvmem} model image. The yellow ellipse shows the size of the natural-weighted beam: $ 0\farcs04 \, \times \, 0\farcs06$. The inset zooms into  the central emission, and the green star marks the star positions. Bottom left panel: 1.65\,$\upmu m$ simulated image of the parametric model. Bottom right panel: 1.3\,mm simulated image of the parametric model. The yellow ellipse shows the size of the  beam: $ 0\farcs04 \, \times \, 0\farcs06$ ***ESTOY CONFUNDIDO POR ESTE BEAM, HAY QUE REPORTAR  1/3 BMAJxBMIN / BPA, INDICANDO 1/3 SEGUIDO POR LOS VALORES EN PESOS NATURALES, Y PLOTEAR EL RESULTADO *****. For all the images in the figure the colour scale is linear.}
  \label{fig:two}
\end{figure*}

Ring13 is surprisingly narrow and seems to be off-centre relative to the GAIA stellar position (at the origin of coordinates in Fig. \ref{fig:two}). We determined the ring's center and orientation by  minimising  the dispersion of the disc radial profiles, from 6\,au to 19\,au, and thus obtained a  PA of 74.6\,$\degr$ ***ADD ERRORES***, an inclination of 33.9\,$\degr$ **ERRORS**** and a centre at $\Delta \alpha = 9\pm0.05$\,mas $\Delta \delta = 0.1\pm0.04$\,mas, relative to the stars.

The ring width can be measured in polar coordinates by fitting 1-D Gaussians, thus  obtaining a width and centroid at each azimuth. On average, we obtained  a FWHM of 4.17$\pm$0.94\,au, and a stellocentric radius of  13.46$\pm$0.43\,au  (See Fig. \ref{fig:polarring}). As the \textsc{uvmem} model image has an effective angular resolution of $\sim$1/3 that of the natural-weighted beam ($0\farcs04 \,\times \, 0\farcs06$) ***CREO QUE AHORA ENTIENDO, HAY QUE ASEGURARTE QUE ESTOS SON LOS BEAMS - DEBIERA SER BMAJORxBMINOR. NO CALZA CON EL ABSTRACT ESO SI***, we see that Ring13 is marginally resolved. After subtaction of the approximate {\tt uvmem} resolution, the ring width is $\sim$*****. 
****SUBTRACT UVMEM BEAM IN QUADRATURE.****

Repeating the optimization of the disk orientation, but  this time aiming for Ring24  with a radiual domain from 20\,au to 70\,au, we obtained a  PA of 76.8\,$\degr$ ***ERROrSS***, with an inclination of 34.0\,$\degr$ **ERRORS* and a centre at $\Delta \alpha = 19\pm0.04$\,mas $\Delta \delta = 7\pm0.03$\,mas relative to the stars. We see that both Ring13 and Ring24  both share / DO NOT SHARE? the same orientation and center, given the errors, and both are offset from the star. Even though all orientation parameters for Ring13 and Ring24  are consistent within the errors, there  are however some hints for a different orientation, as summarised in Fig. \ref{fig:polarring}, perhaps due to the joint effect of all these small differences. ****CALCULATE THE CHI2 DIFFERNCE BETWEEN THE TWO TRACES? SOME FORM OF STATS WOULD BE GOOD****

***I WOULD REMOVE THIS OR IDENTIFY SOMETHING CONCRETE OUT OF IT: We can compare the measured location of the inner ring according to  these parameters versus to the location obtained before. The deviation for each azimuth is shown in Fig. \ref{fig:polarring}, and is not trivial since it depend on a few parameters such as the different inclinations of the rings, the difference in position angle and the shift of the centres relative to the stars. But even though those parameters play an important role, a possible small difference in eccentricities between the inner and the outer ring could also help explain this deviation between the two measurements. ***

@@

The scattered-light image shown in Fig. \ref{fig:two} also shows  a double ring structure in the micron-sized dust distribution. The observed morphology presents an inner cavity of $\sim$10\,au in radius and two rings located at 14.10$\pm$0.01, coincident with Ring13,  and 24.62$\pm$0.08\,au, coincident with Ring24,  with a small gap between them at $\sim$20\,au. The observed second ring  matches   the inner wall of the 1.3\,mm continuum emission outer ring \citep{Ru_z_Rodr_guez_2019}. Two other important features that are present in the image are: the brightness asymmetry and the shadows projected on the disc produced by the close binary system as they eclipse each other, described in detail by \citet{dOrazi}.

\citet{dOrazi} reported a  binary phase in the scattered-light observation around 106\,$\degr$, setting $\phi=$0 from the North and increasing clockwise. Using this measurement, the binary phase was calculated at the time of the ALMA observation at $\phi=$13\,$\degr$. There is no clear signal of any shadow similar to the ones present in the DPI image, this could be explained by innefficient disk cooling compared to the speed of the illumination pattern \citep[][]{Casassus2019MNRAS.486L..58C} ***NO ERA DIFICIL ENCONTRAR ESA REF, PORFA LEER AL MENOS EL ABSTRACT****

\begin{figure}
    \includegraphics[width=\columnwidth]{polar_ring_aprox_and_diff_inner.pdf}
    \caption{(a) Polar decomposition of the 1.3\,mm continuum image, using the orientation of Ring24. We  trace Ring13 using the centroids  (solid line) and width of radial Gaussian fits  (blue region between the dotted lines). (b) Centroid of Ring13, for two disc orientations: the orange line corresponds to the same trace as in a), while the  blue line is  obtained for the inner ring orientation.}
    \label{fig:polarring}
\end{figure}

@@

***HASTA AQUI LLEGUE*****
\section{Parametric radiative transfer model} \label{sec:model}
